---
title: "Using Brasil as VAR lab"
author: "Ricardo Mayer"
date: "October 31, 2017"
output: html_document
---

## Preliminaries
Load all library we will use later, and define any global preferences for this document:


```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, warning=FALSE, message=FALSE}
library(vars)
library(seasonal)
library(forecast)
library(readxl)
library(tidyr)
library(dplyr)
library(timetk)
library(lubridate)
library(ggplot2)
library(stringr)
library(tseries)
library(xts)
library(tibble)
library(tibbletime)
library(tidyquant)
```


## Data import, description and preparation

The initial data is the same currently used by the *Bridge SARIMAX* model (which, out of respect, I will not call it the *BS Model*, but rather the *BRISA* model)

### Data import
In this case, data import is easy because all the heavy work was done by Pablo and quarterly and monthly variables have been nicely saved in a workbook (every country has its own workbook) contaning, among other things, one sheet with quarterly series and another one for monthly series.

Let's take the workbook with data for Brazil and import the two sheets we want

```{r}
bra_q  <- read_excel("data_pablo/Brasil.xlsx", sheet = "quarterly")
bra_m  <- read_excel("data_pablo/Brasil.xlsx", sheet = "monthly")
```

That's it. You can inspect these two data frames (actually, tibbles) by clicking on them or executing `View(mex_q)` in the console

### Data description and exploration
There are several formats appropiate  for our data:
-  First, the *ts* class. Unavoidable if we want to use libraries such as  *forecast* and *vars*
-  But *xts* and *tibbletime* classes are better at managing and transforming data
-  Luckily, the timetk package makes the conversion between these classes very easy to handle. We could in fact have only one version of the data in memory, but for this exploratory stage we will keep xts ts and tibbletime versions of our timeseries data sets.

```{r}
# start and end dates found in the worksheet will come handy
# from tibble to xts

bra_q_tbt <- as_tbl_time(bra_q, index = date)
bra_m_tbt <- as_tbl_time(bra_m, index = date)

bra_q_xts <- tk_xts(bra_q, date_var = date)
bra_m_xts <- tk_xts(bra_m, date_var = date)

# from xts to ts
bra_q_ts <- tk_ts(bra_q_xts,
                 start = c(year(min(index(bra_q_xts))), 
                           quarter(min(index(bra_q_xts)))),
                 frequency = 4)

bra_m_ts <- tk_ts(bra_m_xts,
                 start = c(year(min(index(bra_m_xts))), 
                           quarter(min(index(bra_m_xts)))),
                 frequency = 4)
#store min and max dates
nominal_dates_q <- tibble(name = "q_from_excel",
                     start = min(index(bra_q_xts)),
                     end = max(index(bra_q_xts)))

nominal_dates_m <- tibble(name = "m_from_exel",
                     start = min(index(bra_q_xts)),
                     end = max(index(bra_q_xts)))

```

#### Transform monthly variables into quarterly averages

Since monthly variable are flow variables and not stock, makes more sense to take quarterly averages rather than taking the last monthly value in each quarter, which would be appropriate for stock variables (such as debt)

```{r}
# using xts to convert and then merge
bra_m_xts_to_q <- apply.quarterly(bra_m_xts, mean)
bra_xts <- merge.xts(bra_q_xts, bra_m_xts_to_q)

# using tibbletime to convert and then merge
rolling_mean_na <- rollify(~mean(.x, na.rm = TRUE), window = 3)

bra_m_to_q <- bra_m_tbt %>% 
  tq_transmute(
    mutate_fun =  apply.quarterly,
    FUN = mean,
    na.rm = TRUE
    ) %>% 
  as_tbl_time(index = date)
  

```




